---
title: "Exam_2"
author: "Ricardo Corona"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#clearing the environment
#rm(list=ls())

#loading the inequality dataset
library(rio)
inequality_data = import("inequality.xlsx")

#checking to see if cross sectional or panel
str(inequality_data)
summary(inequality_data)
```

The inequality dataset appears to be a cross sectional dataset as it only incorporates data based on the year 2015. This can be verifies with the above code checking the structure and even recieving a summary of the dataset.

```{r}
#subsetting scores for sweden and denmark
library(dplyr)
gini_scandinavian = subset(inequality_data, country = "Denmark","Sweden")

```

It appears to be better to have lower gini scores overallbased on the previous results

```{r}
head(inequality_data)
#creating a function to remove the accent from belarus
accent.remove = function(x){
  old = "ú"
  new = "u"
  
  x1 = chartr(old , new, x)
}

#running new function on dataset to implemnt change
inequality_data$country = accent.remove(inequality_data$country)

#running head again to ensure change was made
head(inequality_data)

#sorting data by lowest gii inequality scores
inequality_data = inequality_data[order(inequality_data$inequality_gini, decreasing = TRUE),]

#running head again to ensure changes in order were made
head(inequality_data, n = 5)
```

The top 5 countries are Namibia, Zambia, Botswana, Brazil, and Columbia

```{r}
gini_mean = mean(inequality_data$inequality_gini, na.rm = TRUE)
```

The mean iniquality gini score is 36.81 as seen in the code above

```{r}
#creating dummy variables for ineuality using ifelse

ifelse(high_inequality = inequality_data$inequality_gini > 36.81, yes = 1, no = 0)

ifelse(low_inequality = low_inequality < 36.81, yes = 1, no = 0)
```


```{r}
#writing a loop for listing actors
world_actors = c("The World Bank", "The African Development Bank","Bill and Melinda Gates Foundation")

for(i in world_actors){
  print(i)
}
```

```{r}
#importing world bank variable into R
library(WDI)
rural_population = WDI(country = "all",
indicator = c("SP.RUR.TOTL"), 
start = 2015, end = 2015, extra = FALSE, cache = NULL)

#merging variable into inequality dataset
library(tidyverse)
merged_df = left_join(x= inequality_data,
y= rural_population,
by =c("iso2c", "year"))

#removing the extra country.x variable 
merged_df <-
merged_df %>%
select(-c("country.y")) %>% 
rename("country" = "country.x")

#renaming new variable
library(data.table)
setnames(merged_df,"SP.RUR.TOTL", "Rural Population")

#removing NAs on basis of gini and new variable 
na.omit(merged_df, select=c("inequality_gini", "Rural Population"))

#filtering so that we only keep gini scores above 30
data_greater_30 =
merged_df %>%
dplyr::filter(inequality_gini > 30)

#counting how many countries have the sequence "ai" in their name
grep("ai",data_greater_30)


#using the appy function to take the sum of gini
gini_only <- subset(data_greater_30,
select = "inequality_gini")
apply(gini_only, 2, sum)
```

```{r}
#labeling variables in merged_df
library(labelled)
var_label(merged_df) = list(`country` = "Country",
`year` = "year",
`Rural Population` = "Rural population per country",
`iso2c`= "ISO-2 Country Code",
`inequality_gini` = "Gini scores for inequality")

#saving the final dataset
library(rio)
export(merged_df, file = "final_data.dta")
```

